\chapter{Numerical Results}\label{sec:numerical_results}

\section{Introduction}

This chapter presents the results of the numerical experiments conducted to
study the behavioural patterns that can emerge from the interaction between the
two EDs and the EMS.
Additional numerical experiments can also be found in appendix.
%TODO: reference appendix section ~\ref{appendix:more_game_results}.
This chapter consists of the following sections:
\begin{itemize}
    \item Section~\ref{sec:data_collection} describes the data collection
    process.
    \item Section~\ref{sec:dataset_description} gives an overview and some
    descriptive statistics of the data.
    \item Section~\ref{sec:results_what_if} presents the results of the
    numerical experiments and discusses the implications of the results.
\end{itemize}

Note that the data presented in this chapter are archived and can be found
in~\cite{michalis_panayides_2023_7555285}.
This chapter extends the results presented in~\cite{panayides2023game}.


\input{chapters/05_numerical_results/01_data_collection/main.tex}

\input{chapters/05_numerical_results/02_data_description/main.tex}

\input{chapters/05_numerical_results/03_what_if/main.tex}


\section{Chapter summary}

This chapter presents the results of the numerical experiments conducted to
study the behavioural patterns that can emerge from the interaction between the
two EDs and the EMS.
Matrices \(A\), \(B\) (the payoff matrices of the two EDs) and \(R\) (the
routing matrix of the EMS) were generated and analysed for different values of
the parameters to study the behaviour that emerges from playing the game.
Additional numerical experiments can also be found in appendix.
% TODO: reference appendix section ~\ref{appendix:more_game_results}.

Section~\ref{sec:data_collection} describes the data collection process.
The parameter sets that were used to generate the dataset are presented in
along with the structure of the repository that contains all data.
Section~\ref{sec:dataset_description} gives an overview and some descriptive
statistics of the data.
Some plots are also presented to give a visual representation of the data and
to highlight some interesting patterns.

Section~\ref{sec:results_what_if} presents the results of the numerical
experiments and discusses the implications of the results.
More specifically, matrices \(A\), \(B\) and \(R\) are used to run the
asymmetric replicator dynamics algorithm.
Several what-if scenarios are presented and discussed where the system is
flooded with individuals.
Such scenarios include the effect of increasing arrival rates, the effect of
increasing the number of available servers, the effect of increasing the
different capacities and a certain incentivisation approach.
The incentivisation approach is based on the idea carefully penalising some of
the EDs strategies could lead them to choose strategies that are more
cooperative and can help reduce the overall blocking time.
