\chapter{Game theoretic model}\label{sec:game_theoretic_model}


\section{Introduction}

Apart from the queueing theoretic model the second main outcome of this research
is the construction of a game theoretic model that uses the queueing model
described in Section~\ref{sec:queueing_section}.
This game theoretic framework consists of three players where these players (in
Section~\ref{sec:game_ems_ed_application}) will represent the Emergency Medical
Services and two Emergency Departments.
The game theoretic model aims to look into behavioural patterns that emerge
when the players are interacting with each other and act in such a way so that
they maximise their utility. 
This chapter consists of four main sections:

\begin{itemize}
    \item Section~\ref{sec:game_theory_intro} gives a brief introduction to
    the game theoretic concepts
    \item Section~\ref{sec:game_formulation} describes the formulation of the
    game theoretic model that is used in this research
    \item Section~\ref{sec:game_methodology} describes the methodology that is
    used to solve the game theoretic model
    \item Section~\ref{sec:game_ems_ed_application} describes the application
    of the game theoretic model to the Emergency Medical Services and two
    Emergency Departments
\end{itemize}

This chapter extends the concepts described in~\cite{panayides2023game}.


\input{chapters/04_game_theoretic_model/01_game_theory_intro/main.tex}

\input{chapters/04_game_theoretic_model/02_formulation/main.tex}

\input{chapters/04_game_theoretic_model/03_methodology/main.tex}

\input{chapters/04_game_theoretic_model/04_ed_ems_application/main.tex}


\section{Chapter summary}

This chapter introduces a 3-player game theoretic model between the decision 
makers of two queueing systems and a distribution service that distributes
individuals to them.
The game theoretic model is used to investigate the behaviour of the players
when they interact and try to maximise their own utilities.

Section~\ref{sec:game_theory_intro} gives a brief introduction to the game
theoretic concepts that are used in this chapter.
A brief introduction on Normal Form games and the concept of Nash equilibrium
are given, along with some examples of games and their equivalent Nash
equilibria.
Additionally, some learning algorithms are introduced.
The learning algorithm that is mostly used in this chapter is the
asymmetric replicator dynamics algorithm.
Finally, a description of perfect-information and imperfect-information normal
form games is given.

Section~\ref{sec:game_formulation} describes the formulation of the game
theoretic model that is used in this thesis.
An overview of the three players and the parameters that are used in the game
is given.
The parameter that is of most interest here is the threshold parameter that
is used to determine when individuals are being blocked from entering node
\(1\) of the queueing system.
The set of strategies of the two players is essentially all possible values
that the threshold parameter can take.
Furthermore, the payoffs for the two players are described.
The distribution service's payoff is determined by the number of individuals
that are blocked in node \(2\) of the queueing system.
The queueing system's payoff is determined by the proportion of individuals
whose waiting time is less than a predefined target time.
Thus, the imperfect-information extensive form game is introduced.

Section~\ref{sec:game_methodology} describes the methodology that is used to
solve the game theoretic model.
The methodology uses Brent's algorithm to find
the optimal split of individuals that the distribution service can distribute
to the two queueing systems to minimise its own blocking time.
The routing matrix is then described that contains all possible values of the
proportion of individuals that can be distributed to the two queueing systems.
The game theoretic model is then reduced to a 2-player normal form game where
the utilities of the game are decided the distribution service.
Following this, a game with some example set of parameters is given and
is solved using the support enumeration algorithm and the Lemke-Howson
algorithm.
Additionally, some learning algorithms are applied to the game to observe
the behaviour of the players when they interact together.
The learning algorithms that are used are fictitious play, stochastic
fictitious play, and asymmetric replicator dynamics.

Finally, Section~\ref{sec:game_ems_ed_application} maps the game theoretic
model to a healthcare setting to observe the behavioural gaming that takes place
at the EMS~-~ED interface.
The three players now become the Emergency Medical Services (EMS) and two
Emergency Departments (EDs).
The game essentially consists of the EMS deciding what proportion of
ambulances to send to each hospital so that the blocking time is minimised and
the EDs choosing what threshold parameter to use so that a certain policy
objective is met.
In line with the current NHS policy~\cite{fourhourtarget},
the objective is set to be that \(95\%\) of the patients that are admitted to
the EDs should be admitted within \(4\) hours.
