\chapter{The \texttt{ambulance\_game} Python library}
\label{appendix:ambulance_game}

% TODO: Reference sections of thesis throughout
This chapter of the appendix provides an overview of the
\texttt{ambulance\_game} Python library.
This is a library that accompanies all chapters of this thesis and provides
the functionality to run the mathematical models and simulations described
in the thesis.
The library uses the best available tools to ensure the code is correct,
readable, well-documented and properly tested.
These tools are outlined in Section~\ref{sec:intro_test_code}.
The library is also publicly available on GitHub and has been archived on
Zenodo~\cite{ambulance_game_github_repo}.

This appendix chapter is structured based on the Diataxis
framework~\cite{Procida_Diataxis_documentation_framework}.
The structure of the chapter is as follows:

\begin{itemize}
    \item Section~\ref{sec:ambulance_game_installation} provides instructions
        on how to install the library.
    \item Section~\ref{sec:ambulance_game_tutorial} a learning-oriented lesson
        on performing a specific task using the library.
    \item Section~\ref{sec:ambulance_game_how_to} provides a goal-oriented
        how-to guide on how to use the library.
    \item Section~\ref{sec:ambulance_game_reference} provides some technical
        descriptions of the library and how to use it.
    \item Section~\ref{sec:ambulance_game_explanation} provides a discussion
        of some of the details of the library.
\end{itemize}


\input{chapters/00_appendix/01_ambulance_game_library/01_installation/main.tex}

\input{chapters/00_appendix/01_ambulance_game_library/02_tutorial/main.tex}

\input{chapters/00_appendix/01_ambulance_game_library/03_how_to_guides/main.tex}

\input{chapters/00_appendix/01_ambulance_game_library/04_reference/main.tex}

\input{chapters/00_appendix/01_ambulance_game_library/05_explanation/main.tex}
