\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}
% TODO: Write more here?
First of all I would like to express my sincerest and greatest gratitude to my
supervisors, Dr. Vince Knight and Prof. Paul Harper whose guidance and support
was the most helpful part throughout my PhD.
Their tolerance, and great sense of humour, helped me through the ups and downs
that come with a PhD and I am very grateful for that.
I truly believe I could not have had a better pair of supervisors.
Thanks to the both of you.

I would also like to thank The Healthcare Improvement Studies Institute (THIS)
for funding this research and for providing me with this wonderful experience
and Prof. Davina Allen for her comments and ethnographic insights that she
provided that helped guide my research.

% TODO: Sonya Crowe?
% In particular I would like to thank my THIS mentor Prof. Sonya Crowe for her
% advice and support throughout my PhD.
% Those couple of meetings were really helpful and valuable.

On a more personal note I would like to thank my parents, Anastasia and
Panayiotis and my brother Stavros for their unconditional love and support
throughout my PhD.
% TODO: Rewrite the next sentence
A big thank you to my friends in Cardiff as well; my flatmates Nikolas and
Vasilis for the necessary breaks; Athena for the countless Thursday lunch
breaks; and Panos for still trying to teach me what goes to which bin.

I would also like to thank all of my academic buddies and in particular
Elizabeth Williams for her constant help with any random problem I had through
my PhD and for looking out for MY plant Trev; Matthew Howells for never failing
to make the right pun and for the countless nerf gun battles we had; and
Michela Corradini for being the most annoying individual I ever met, and for
that one thing she helped me fix on Figure~\ref{fig:venn_diagram_with_subsets}
in page~\pageref{fig:venn_diagram_with_subsets} of this thesis and kept saying
I should include her in my acknowledgements for it.
Lastly, thank you to the School of Mathematics staff for maintaining the
friendly and supportive environment that I have enjoyed throughout my PhD.
