\chapter*{Dissemination of Work}
\addcontentsline{toc}{chapter}{Dissemination of Work}

\section*{Publications}
\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
2022: \textbf{A game theoretic model of the behavioural gaming that takes place
at the EMS - ED interface.} Michalis Panayides, Vince Knight, Paul Harper -
\href{https://www.sciencedirect.com/science/article/pii/S0377221722005549?via%3Dihub}{ScienceDirect} -
\end{enumerate}

\section*{Talks \& Posters}

All copies of the slides and posters are publicly available on
\href{https://github.com/MichalisPanayides/Talks}{GitHub}.

\textsc{Talks}
\begin{itemize}
    \item \textbf{A game theoretic model of the behavioural gaming that takes
    place at the EMS-ED interface.} \emph{THIS Space Event, 2020}
    \item \textbf{Using Python to measure the expected wait in a queue with two
    waiting rooms.} \emph{PyCon Namibia, 2020}
    \item \textbf{How to make an awesome package in python.} \emph{SIAM/IMA PGR
    seminar series, 2021}
    \item \textbf{The Ambulance Decision Game.} \emph{Poster for School of
    Mathematics poster day, 2021}
    \item \textbf{A 3-player game theoretic model of a choice between two
    queueing systems with strategic managerial decision making.}
    \emph{EURO, 2021}
    \item \textbf{A 3-player game theoretic model of a choice between two
    queueing systems with strategic managerial decision making.}
    \emph{Wales mathematics Colloquium, Gregynog Hall, 2022}
    \item \textbf{Recovering from inefficiencies in queueing systems with two
    consecutive waiting zones.} \emph{CORS/INFORMS conference, 2022}
    \item \textbf{A game theoretic model between two Emergency Departments and
    the Emergency Medical Services.}
    \emph{ORAHS/CHOIR seminar series, University of Twente, 2022}
    \item \textbf{A 3-player game theoretic model of a choice between two
    queueing systems with strategic managerial decision making.}
    \emph{EURO, 2022}
    \item \textbf{A game theoretic model between two Emergency Departments and
    the Emergency Medical Services.} \emph{SWORDS, 2022}
    \item \textbf{Playing games - An introduction to game theory.}
    \emph{Cardiff University, School outreach day, 2023}
\end{itemize}

\section*{Software development}

\begin{itemize}
\item
\href{https://github.com/MichalisPanayides/AmbulanceDecisionGame}{Ambulance\_game},
an open source python library for the development of a game theoretic model
between two Emergency Departments and the Emergency Medical Services.
\textbf{Contribution:} Main developer.

\item \href{https://github.com/CiwPython/Ciw}{Ciw}, a discrete event simulation
python library for open queueing networks.
\textbf{Contributions:} Implementation of server dependent distributions,
implementation of server priorities.

\item \href{https://github.com/drvinceknight/Nashpy}{Nashpy}, a Python library
used for the computation of equilibria in 2 player strategic form games.
\textbf{Contributions:} Implementation of the learning algorithm, asymmetric
replicator dynamics.

\end{itemize}